---
title             : "Appendix 1: Plots of Simulation Results"
shorttitle        : "2S-PA-Int"

author: 
  - name          : "Gengrui (Jimmy) Zhang"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    email         : "gengruiz@email.com"
    role: # Contributorship roles (e.g., CRediT, https://credit.niso.org/)
      - "Conceptualization"
      - "Writing - Original Draft Preparation"
      - "Writing - Review & Editing"

affiliation:
  - id            : "1"
    institution   : "University of Southhern California"

floatsintext      : no
linenumbers       : yes
draft             : no
mask              : no

figurelist        : no
tablelist         : no
footnotelist      : no

classoption       : "man"
csl               : "`r system.file('rmd', 'apa7.csl', package = 'papaja')`"
output            : papaja::apa6_pdf
---

```{r setup, include = FALSE}
library("papaja")
library(haven)
library(dplyr)
library(here)
library(tidyverse)
```

```{r message=FALSE, warning=FALSE}
sim <- readRDS(here("Sim_Data", "continuous_bound_09252024.rds"))
sim_results <- sim %>% 
  dplyr::select(-c(REPLICATIONS:ERRORS)) %>%
  mutate(across(where(is.numeric), round, 4),
         cor_xm = ifelse(cor_xm == 0, "0", cor_xm))
sim_null <- sim_results %>% filter(gamma_xm == 0)
sim_alt <- sim_results %>% filter(gamma_xm != 0)
```

Four plots were provided to visualize the pattern of standardized bias, relative standard error (SE) bias, coverage rate of 95% CI, and empirical type I error rate.

```{r standardized bias plot, message=FALSE, warning=FALSE}
# Helper Function 1
process_data <- function(data) {
  results <- data %>%
    gather("var", "val", raw_bias.mmr.est:rmse.tspa.est) %>%
    dplyr::select(-c(warning_total.mmr.warnings_count:WARNINGS)) %>%
    separate(col = var, into = c("stats", "method"), sep = "\\.") %>%
    spread(stats, val) %>%
    mutate(N_lab = as_factor(paste0("italic(N) == ", N)),
           gammaint_lab = as_factor(paste0("\\beta_{xm} == ", gamma_xm)),
           cor_xm_lab = as_factor(paste0("corr[XM] == ", cor_xm)),
           rel_lab = as_factor(paste0("rho == ", rel)))

  return(results)
}

# Helper Function 2
plot_results <- function(data,
                         x_var, y_var, 
                         x_label, y_label, 
                         y_limits = NULL, 
                         color_var = "method", 
                         shape_var = "method", 
                         point_size = 3,         # Increased point size
                         point_alpha = 0.9,      # Added point alpha for more flexibility
                         line_width = 1,         # Line width for lines if `add_lines = TRUE`
                         hollow_shapes = FALSE,  # Option for hollow shapes
                         shape_options = NULL,   # New option for custom shapes
                         color_options = NULL,   # New option for custom colors
                         add_lines = FALSE, 
                         remove_baseline = FALSE) {
  
  # Optionally filter out "mmr" group
  if (remove_baseline) {
    data <- data %>% filter(!!sym(color_var) != "mmr")
  }
  
  # Default shapes if `shape_options` is not provided
  default_shapes <- if (hollow_shapes) {
    if (remove_baseline) {
      c("rapi" = 21, "upi" = 24, "tspa" = 22)  # Hollow shapes 21-25
    } else {
      c("mmr" = 23, "rapi" = 21, "upi" = 24, "tspa" = 22)
    }
  } else {
    if (remove_baseline) {
      c("rapi" = 16, "upi" = 17, "tspa" = 15)  # Filled shapes
    } else {
      c("mmr" = 18, "rapi" = 16, "upi" = 17, "tspa" = 15)
    }
  }
  
  # Default colors if `color_options` is not provided
  default_colors <- if (remove_baseline) {
    c("rapi" = "#E69F00", "upi" = "#CC79A7", "tspa" = "#009E73")
  } else {
    c("mmr" = "pink", "rapi" = "#CC79A7", "upi" = "#d95f02", "tspa" = "#009E73")
  }
  
  # Use `shape_options` if provided, otherwise use defaults
  shapes <- if (!is.null(shape_options)) shape_options else default_shapes
  
  # Use `color_options` if provided, otherwise use defaults
  colors <- if (!is.null(color_options)) color_options else default_colors

  plot <- data %>%
    ggplot(aes(x = factor(!!sym(x_var)), y = !!sym(y_var), color = !!sym(color_var), shape = !!sym(shape_var))) +
    geom_point(position = position_jitterdodge(jitter.width = 0, dodge.width = 0), 
               size = point_size, alpha = point_alpha, 
               fill = if (hollow_shapes) "white" else NA) +  # Fill color for hollow shapes
    facet_grid(cor_xm_lab ~ rel_lab, labeller = label_parsed) +
    scale_shape_manual(name = "Method", values = shapes,
                       labels = if (remove_baseline) c("rapi" = "RAPI", "upi" = "Matched-Pair UPI", "tspa" = "2S-PA-Int") else c("mmr" = "MMR", "rapi" = "RAPI", "upi" = "Matched-Pair UPI", "tspa" = "2S-PA-Int")) +
    scale_color_manual(name = "Method", values = colors,
                       labels = if (remove_baseline) c("rapi" = "RAPI", "upi" = "Matched-Pair UPI", "tspa" = "2S-PA-Int") else c("mmr" = "MMR", "rapi" = "RAPI", "upi" = "Matched-Pair UPI", "tspa" = "2S-PA-Int")) +
    labs(x = x_label, y = y_label, color = "Method", shape = "Method") +
    theme_minimal() +
    theme(legend.title = element_text(size = 12, face = "bold"),
          legend.text = element_text(size = 10),
          legend.position = "right",
          strip.text = element_text(size = 10, face = "bold"),  # Adjust facet label text
          panel.border = element_rect(color = "black", fill = NA, size = 1),  # Add black border to panels
          strip.background = element_rect(color = "black", size = 1),  # Add black border to facet labels
          panel.grid.major = element_line(size = 0.5, linetype = 'solid', colour = "gray80"),
          panel.grid.minor = element_blank(),
          panel.grid.major.x = element_blank(),  # Remove vertical grid lines
          panel.grid.major.y = element_line(size = 0.5, linetype = 'solid', colour = "gray80"),  # Keep horizontal lines
          axis.line = element_line(color = "black"),  # Add black lines to x and y axes
          axis.ticks = element_line(color = "black"))
  
  if (add_lines) {
    plot <- plot + 
      geom_line(aes(group = !!sym(color_var), 
                    linetype = !!sym(shape_var)), 
                size = line_width, alpha = 0.6) +
      scale_linetype_manual(name = "Method",
                            values = if (remove_baseline) c("rapi" = "solid", "upi" = "dashed", "tspa" = "twodash") else c("mmr" = "dotted", "rapi" = "solid", "upi" = "dashed", "tspa" = "twodash"),
                            labels = if (remove_baseline) c("rapi" = "RAPI", "upi" = "Matched-Pair UPI", "tspa" = "2S-PA-Int") else c("mmr" = "MMR", "rapi" = "RAPI", "upi" = "Matched-Pair UPI", "tspa" = "2S-PA-Int"))
  }
  
  if (!is.null(y_limits)) {
    plot <- plot + ylim(y_limits)
  }
  
  return(plot)
}

# Plot
type1_pd <- process_data(sim_null)
power_pd <- process_data(sim_alt)

custom_colors <- c("mmr" = "pink", "rapi" = "#CC79A7", "upi" = "#d95f02", "tspa" = "#009E73")
custom_shapes <- c("mmr" = 23, "rapi" = 21, "upi" = 3, "tspa" = 24)

sd_bias_plot <- plot_results(power_pd,
                             x_var = "N",
                             y_var = "std_bias",
                             x_label = "Sample Size (N)",
                             y_label = "Standardized Bias",
                             y_limits = c(-0.45, 0.45),
                             point_size = 1.5,
                             point_alpha = 3,
                             line_width = 0.5,
                             hollow_shapes = TRUE,          # Enable hollow shapes
                             shape_options = custom_shapes, # Custom shapes
                             color_options = custom_colors, # Custom colors
                             add_lines = FALSE,
                             remove_baseline = FALSE)

sd_bias_plot <- sd_bias_plot +
  geom_hline(yintercept = c(-0.4, 0.4), linetype = "dashed", color = "red")

ggsave("/Users/jimmy_z/R Projects/2SPA-Int/Paper/Qual_1_Supplemental_Material/2S_PA_Int_Plot_of_Simulation/sd_bias_plot.png", plot = sd_bias_plot, width = 7, height = 5, dpi = 300)
```

```{r RSE bias (plot), message=FALSE, warning=FALSE}
rse_bias_plot <- plot_results(power_pd,
                             x_var = "N",
                             y_var = "stdMed_rse_bias",
                             x_label = "Sample Size (N)",
                             y_label = "Relative SE Bias",
                             y_limits = c(-0.25, 0.25),
                             point_size = 1.5,
                             point_alpha = 3,
                             line_width = 0.5,
                             hollow_shapes = TRUE,          # Enable hollow shapes
                             shape_options = custom_shapes, # Custom shapes
                             color_options = custom_colors, # Custom colors
                             add_lines = FALSE,
                             remove_baseline = FALSE)
rse_bias_plot <- rse_bias_plot +
  geom_hline(yintercept = c(-0.1, 0.1), linetype = "dashed", color = "red")

ggsave("/Users/jimmy_z/R Projects/2SPA-Int/Paper/Qual_1_Supplemental_Material/2S_PA_Int_Plot_of_Simulation/rse_bias_plot.png", plot = rse_bias_plot, width = 7, height = 5, dpi = 300)
```

```{r coverage (plot), message=FALSE, warning=FALSE}
cov_plot <- plot_results(power_pd,
                         x_var = "N",
                         y_var = "coverage_std",
                         x_label = "Sample Size (N)",
                         y_label = "Coverage Rate of 95% CI",
                         y_limits = c(0.8, 1),
                         point_size = 1.5,
                         point_alpha = 3,
                         line_width = 0.5,
                         hollow_shapes = TRUE,
                         shape_options = custom_shapes, # Custom shapes
                         color_options = custom_colors, # Custom colors
                         add_lines = FALSE,
                         remove_baseline = FALSE)
cov_plot <- cov_plot +
  geom_hline(yintercept = c(0.95), linetype = "dashed", color = "red")

ggsave("/Users/jimmy_z/R Projects/2SPA-Int/Paper/Qual_1_Supplemental_Material/2S_PA_Int_Plot_of_Simulation/cov_plot.png", plot = cov_plot, width = 7, height = 5, dpi = 300)
```

```{r Type I Error (plot), message=FALSE, warning=FALSE}
type1_plot <- plot_results(type1_pd,
                           x_var = "N",
                           y_var = "type1_lrt",
                           x_label = "Sample Size (N)",
                           y_label = "Type I Error Rate",
                           y_limits = c(0, 0.1),
                           point_size = 1.5,
                           point_alpha = 3,
                           line_width = 0.5,
                           hollow_shapes = TRUE,
                           shape_options = custom_shapes, # Custom shapes
                           color_options = custom_colors, # Custom colors
                           add_lines = FALSE,
                           remove_baseline = FALSE)

ggsave("/Users/jimmy_z/R Projects/2SPA-Int/Paper/Qual_1_Supplemental_Material/2S_PA_Int_Plot_of_Simulation/type1_plot.png", plot = type1_plot, width = 7, height = 5, dpi = 300)
```

\begin{figure}[h]
\centering
\includegraphics[width=1.2\textwidth]{/Users/jimmy_z/R Projects/2SPA-Int/Paper/Qual_1_Supplemental_Material/2S_PA_Int_Plot_of_Simulation/sd_bias_plot.png}
\caption{Standardized Bias in Latent Interaction Effect Estimates for Nonzero Effect}
\caption*{\textit{Note. Points with different shapes represent values for the three latent interaction methods across all simulation conditions. Some values of MMR fall outside the y-axis limits and are not shown. The red dashed lines indicate the acceptable range of [-0.40, 0.40].}}
\end{figure}

\begin{figure}[h]
\centering
\includegraphics[width=1.2\textwidth]{/Users/jimmy_z/R Projects/2SPA-Int/Paper/Qual_1_Supplemental_Material/2S_PA_Int_Plot_of_Simulation/rse_bias_plot.png}
\caption{Relative SE Bias in Latent Interaction Effect Estimates for Nonzero Effect}
\caption*{\textit{Note. Points with different shapes represent values for the three latent interaction methods across all simulation conditions. The red dashed lines indicate the acceptable range of [-10\%, 10\%].}}
\end{figure}

\begin{figure}[h]
\centering
\includegraphics[width=1.2\textwidth]{/Users/jimmy_z/R Projects/2SPA-Int/Paper/Qual_1_Supplemental_Material/2S_PA_Int_Plot_of_Simulation/cov_plot.png}
\caption{Coverage Rate of 95\%CI in Latent Interaction Effect Estimates for Nonzero Effect}
\caption*{\textit{Note. Points with various shapes represent values of three latent interaction methods for all simulation conditions. The red dashed lines indicate the level of 95\%.}}
\end{figure}

\begin{figure}[h]
\centering
\includegraphics[width=1.2\textwidth]{/Users/jimmy_z/R Projects/2SPA-Int/Paper/Qual_1_Supplemental_Material/2S_PA_Int_Plot_of_Simulation/type1_plot.png}
\caption{Empirical Type I Error Rates in Latent Interaction Effect Estimates for Zero Effect}
\caption*{\textit{Note. Points with various shapes represent values of three latent interaction methods for all simulation conditions.}}
\end{figure}
